What can Issues on GitHub and Google Code tell us about Bug Hazards in Android and Java?

### 2015-02-12 android-stacks.tex
section 4 - changes

The goal of this exploratory mining study was to investigate whether the exception stack traces 
embedded on issues of Android projects (hosted in GitHub and Google Code), can reveal 
\emph{bug hazards} on the exception-related code of both the applications and the underlying framework. 
This exploration is guided by the set of best practices compiled in this study; we investigate whether characteristics 
of exception stack traces and the related source code it can pinpoint best practices violations and therefore to bug hazards.
As mentioned before, in this context ``bug hazards" are the characteristics on the exception-related code 
that favor the introduction of failures such as uncaught exceptions.

@@

Our study focused on open-source apps, since the information needed to perform our study
cannot be retrieved from commercial apps, whose issue report systems and 
source codes are not publicly available. The Android open-source apps were also 
target of other research studies ~\cite{Linar13,ahimed}.

In this study we explore the domain quantitatively and highlight interesting cases by 
exploring cases qualitatively.  Figure~\ref{fig:overview} gives an overview of the study, whose tasks are detailed next.

\subsection{Data Extraction Process}
\label{sec:miningproc}

In this exploratory study the following information is needed: (i) the issues reported on Android project hosted on 
GitHub and Google Code; (ii) the stack traces embedded on these issues; (iii) the type of the exceptions
 reported on the stack traces (e.g., error, runtime, checked); (iv) the origin of such exceptions 
(e.g., the application, a library, the Android platform). Next, we describe: how such information 
was retrieved from GitHub and Google Code; how the exceptions stack traces were extracted and distilled,
 and how the type associated to each exception and its origins was identified, and finally how manual
inspections supported this work.

@@

\subsubsection{Android Apps in GitHub}
\label{sec:git}

@@

Then we performed further cleanup, inspecting the site of every Android
reporting at least one stack trace, to make sure that they represented real
mobile apps. During this clean up 106 apps were removed because they were either
example projects (i.e., toy projects) or tools to support Android development
(e.g. Selendroid, Roboeletric - tools to support the testing of Android apps).
The filtered set consisted of 482 apps. This set of 482 projects contained overall 31,592 issues from which 4,042 exception stack traces 
were extracted. Issues on Github are different from issues on dedicated bug tracking tools such as 
Bugzilla and Jira. The most important difference is that there are no predefined fields
  (e.g. severity and priority). Instead, Github uses a more open ended tagging system, on which
repositories are offered a pre-defined set of labels, but repository owners can modify 
them at will. Therefore, an issue may have none or an arbitrary set of labels depending 
on which repository it was created. Table~\ref{tab:lables} illustrates the ocurrences of different labels 
on the issues including exception stack traces.

Based on the assumption that regardless the issue type, every exception stack
trace contains relevant information concerning the exception structure of the
projects analyzed - and therefore can review bug hazards on the exception-related code -  
we opted for not restricting the analysis only on defect
issues\footnote{We conducted the same analysis on the defect issues and the top
exceptions are similar to the ones mentioned on defect issues. Due to space
limitation we limit to present the general analysis here. More detailed analysis
on the defect issues can be found at:
\url{https://github.com/souzacoelho/exceptionminer}}.
 Moreover, issues and pull requests are dual on Github; all pull requests have a corresponding 
``backing'' issue (i.e., an issue that is automatically generated per pull request). Therefore, we excluded such automatically generated
issues from our analysis.

\subsubsection{Android Apps in Google Code}
Google Code contains largely used open-source Android apps (e.g. K9Mail \footnote{K9Mail moved to Github but as a way of not loosing the project history 
it advises their users to report bugs on Google Code issue tracker: https://github.com/k9mail/k-9/wiki/LoggingErrors.}),
 however, differently from GitHub, Google Code does not provide an API to access the information related
 to hosted projects \footnote{Google code used to provided a Web service to such repository, but it was deactivated in June 2013 what Google called a `clean-up action".}.
To overcome this limitation we needed to implement a Web Crawler (incorporated in ExceptionMiner tool described next) that navigated 
through the web interface of Google Code projects extracting all issues and issue comments and storing in a relational database for later analysis.

@@

\section{The ExceptionMiner Tool}
\label{sec:exceptionminer}

@@

\emph{\textbf{Exception Stack Trace Distiller.}}
This component distills the information that composes a stack trace.
 Some of the attributes extracted from the stack trace are:
 the root exception and its signaler, the exception wrappers and their corresponding signalers. 
This component also distills fine grained information of each attribute such as: the classes and packages associated to them.
This component is based on a combination between a regular expression based parser 
with heuristics able to identify and filter exception names and stack traces inline with text. 
In contrast to existing issue parsing solutions such as Infozilla, the parser
created in this work can discover stack traces inlined with logs files \footnote{In several 
exception stack traces, the exception frames were preceeded by logging information e.g., 
"03-01 15:55:01.609 (7924): at android.app.ActivityThread.access\$600(ActivityThread.java:127) 
"which could not be detected by existing tools.)}.

>>>>

The goal of this exploratory mining study was to investigate whether the exception stack traces 
embedded on issues of Android projects (hosted in GitHub and Google Code), can reveal 
\emph{bug hazards} on the exception-related code of both the applications and the underlying framework. 
This exploration is guided by the set of best practices compiled in this study. We investigate whether characteristics 
of exception stack traces can pinpoint best practices violations and therefore reveal bug hazards.
As mentioned before, in this context ``bug hazards" are the characteristics on the exception-related code 
that favor the introduction of failures such as uncaught exceptions.

@@

Figure~\ref{fig:overview} gives an overview of the study.
Firstly  the issues reported on Android projects hosted on GitHub and Google Code were recovered (Figure~\ref{fig:overview} - (1) and (2)).
Then the stack traces embedded each issue were extracted and distilled (Figure~\ref{fig:overview} - (3)).
The stack trace information were then combined with source code and bytecode analysis in order to discover 
the type of the exceptions reported on the stack traces (e.g., error, runtime, checked)  (Figure~\ref{fig:overview} - (5))  and the origin of such exceptions 
(e.g., the application, a library, the Android platform)  (Figure~\ref{fig:overview} - (6)). 
Manual inspections were also used to support this mining process (Figure~\ref{fig:overview} - (4), (7) and (9)) and support
the search for bug hazards performed in this study  (Figure~\ref{fig:overview} - (8)). Next sections detail each step of this mining process.

Our study focused on open-source apps, since the information needed to perform our study
cannot be retrieved from commercial apps, whose issue report systems and 
source codes are not publicly available. The Android open-source apps were also 
target of other research studies ~\cite{Linar13,ahimed}.   
In this study we explore the domain quantitatively and highlight interesting cases by 
exploring cases qualitatively.

%\subsection{Data Extraction Process}
%\label{sec:miningproc}

%In this exploratory study the following information is needed: (i) the issues reported on Android projects hosted on 
%GitHub and Google Code; (ii) the stack traces embedded on these issues; (iii) the type of the exceptions
 %reported on the stack traces (e.g., error, runtime, checked); (iv) the origin of such exceptions 
%(e.g., the application, a library, the Android platform). Next, we describe: how such information 
%was retrieved from GitHub and Google Code.

@@

\subsection{Android Apps in GitHub}
\label{sec:git}

@@

Then we performed further cleanup, inspecting the site of every Android project
reporting at least one stack trace, to make sure that they represented real
mobile apps. During this clean up 106 apps were removed because they were either
example projects (i.e., toy projects) or tools to support Android development
(e.g. Selendroid, Roboeletric - tools to support the testing of Android apps).
The filtered set consisted of 482 apps. This set of 482 projects contained overall 31,592 issues from which 4,042 exception stack traces 
were extracted. Issues on Github are different from issues on dedicated bug tracking tools such as 
Bugzilla and Jira. The most important difference is that there are no predefined fields
  (e.g. severity and priority). Instead, Github uses a more open ended tagging system, on which
repositories are offered a pre-defined set of labels, but repository owners can modify 
them at will. Therefore, an issue may have none or an arbitrary set of labels depending 
on which repository it was created. Table~\ref{tab:lables} illustrates the ocurrences of different labels 
on the issues including exception stack traces.

Based on the assumption that regardless the issue lables, every exception stack
trace contains relevant information concerning the exception structure of the
projects analyzed - and therefore can review bug hazards on the exception-related code -  
we opted for not restricting the analysis only on defect issues\footnote{We conducted the same analysis 
on the defect issues only and the top exceptions were quite similar.
 Due to space limitation we limit to present the general analysis here. More detailed analysis
on the defect issues can be found at:\url{https://github.com/souzacoelho/exceptionminer}}.

\subsection{Android Apps in Google Code}
Google Code contains largely used open-source Android apps (e.g. K9Mail \footnote{K9Mail moved to Github but as a way of not loosing the project history 
it advises their users to report bugs on Google Code issue tracker: https://github.com/k9mail/k-9/wiki/LoggingErrors.}),
 however, differently from GitHub, Google Code does not provide an API to access the information related
 to hosted projects \footnote{Google code used to provided a Web service to such repository, but it was deactivated in June 2013 what Google called a `clean-up action".}.
To overcome this limitation we needed to implement a Web Crawler (incorporated in ExceptionMiner tool described next) that navigated 
through the web interface of Google Code projects extracting all issues and issue comments and storing in a relational database for later analysis.

@@

\subsection{The ExceptionMiner Tool}
\label{sec:exceptionminer}

@@

\emph{\textbf{Exception Stack Trace Distiller.}}
This component distills the information that composes a stack trace.
 Some of the attributes extracted from the stack trace are:
 the root exception and its signaler, the exception wrappers and their corresponding signalers. 
This component also distills fine grained information of each attribute such as: the classes and packages associated to them.
This component is based on a combination between a parser (based on regular expressions) 
and heuristics able to identify and filter exception names and stack traces inline with text. 
In contrast to existing issue parsing solutions such as Infozilla, the parser
created in this work can discover stack traces mixed with log information\footnote{In several 
exception stack traces, the exception frames were preceeded by logging information e.g., 
"03-01 15:55:01.609 (7924): at android.app.ActivityThread.access\$600(ActivityThread.java:127) 
"which could not be detected by existing tools.)}.
